---
title: "Assembling and cleaning full SRA metadata"
output: html_notebook
---

```{r packages, message=F, warning=F}
library(tidyverse); library(rjson); library(vroom)

# Import several functions and constants for cleaning
source("./annotation_cleaning_functions.R")
source("./non_meaningful_disease_annotations.R")
source("./real_study_diseases.R")
source("./ethnicities_list.R")
source("./meaningful_treatments.R")
```

# Assembling raw data

#### Load raw data
```{r}
df.original_metadata <- vroom("./sra_compiled_metadata2020_10_06_185340.csv") %>% 
  mutate_all(tolower) %>% 
  mutate_all(trimws) 

df.values <- read_csv("all_value_annotations_uncleaned.csv")%>% 
  mutate_all(trimws) %>% 
  mutate_all(function(x) ifelse(x=="null", NA, x))

df.studies <- read_csv("all_study_title_annotations_uncleaned.csv")%>% 
  mutate_all(trimws) %>% 
  mutate_all(function(x) ifelse(x=="null", NA, x))
```

#### Compiling uncleaned metadata annotations for values and study titles
```{r Compile values, message=F, warning=F}
## Original code for converting MetaAnnotator JSON data to a compiled dataframe

# mondo_1 <- "ontology_annotations/allValues_MONDO/"
# mondo_2 <- "ontology_annotations/missingValues_MONDO/"
# meddra_1 <- "ontology_annotations/filteredValues_MEDDRA/"
# meddra_2 <- "ontology_annotations/missingValues_MEDDRA/"
# bto_1 <- "ontology_annotations/filteredValues_BTO/"
# bto_2 <- "ontology_annotations/missingValues_BTO/"
# clo_1 <- "ontology_annotations/filteredValues_CLO/"
# clo_2 <- "ontology_annotations/missingValues_CLO/"
# uberon_1 <- "ontology_annotations/filteredValues_UBERON/"
# uberon_2 <- "ontology_annotations/missingValues_UBERON/"
# dron_1 <- "ontology_annotations/filteredValues_DRON/"
# dron_2 <- "ontology_annotations/missingValues_DRON/"
# 
# df.original <- create_annotationDF_fromDir(mondo_1) %>% rename(value = attributeName, value_MONDO = attributeName_annotated) %>%
#   left_join(create_annotationDF_fromDir(bto_1) %>% 
#               rename(value = attributeName, value_BTO = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(uberon_1) %>% 
#               rename(value = attributeName, value_UBERON = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(meddra_1) %>% 
#               rename(value = attributeName, value_MEDDRA = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(clo_1) %>% 
#               rename(value = attributeName, value_CLO = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(dron_1) %>% 
#               rename(value = attributeName, value_DRON = attributeName_annotated), by = "value")
# 
# df.missing <- create_annotationDF_fromDir(mondo_2) %>% rename(value = attributeName, value_MONDO = attributeName_annotated) %>%
#   left_join(create_annotationDF_fromDir(bto_2) %>% 
#               rename(value = attributeName, value_BTO = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(uberon_2) %>% 
#               rename(value = attributeName, value_UBERON = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(meddra_2) %>% 
#               rename(value = attributeName, value_MEDDRA = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(clo_2) %>% 
#               rename(value = attributeName, value_CLO = attributeName_annotated), by = "value") %>%
#   left_join(create_annotationDF_fromDir(dron_2) %>% 
#               rename(value = attributeName, value_DRON = attributeName_annotated), by = "value")
# 
# df <- df.original %>% bind_rows(df.missing) %>% mutate_all(tolower) %>% distinct()
# 
# write_csv(df, "all_value_annotations_uncleaned.csv")
```

```{r Compile study titles, message=F, warning=F}
# # Original code for converting MetaAnnotator JSON data to a compiled dataframe
# 
# mondo_1 <- "ontology_annotations/studyTitle_MONDO/"
# mondo_2 <- "ontology_annotations/missing_studyTitle_MONDO/"
# meddra_1 <- "ontology_annotations/studyTitle_MEDDRA/"
# meddra_2 <- "ontology_annotations/missing_studyTitle_MEDDRA/"
# bto_1 <- "ontology_annotations/studyTitle_BTO/"
# bto_2 <- "ontology_annotations/missing_studyTitle_BTO/"
# uberon_1 <- "ontology_annotations/studyTitle_UBERON/"
# uberon_2 <- "ontology_annotations/missing_studyTitle_UBERON/"
# dron_1 <- "ontology_annotations/studyTitle_DRON/"
# dron_2 <- "ontology_annotations/missing_studyTitle_DRON/"
# 
# df.study <- create_annotationDF_fromDir(mondo_1) %>% rename(study_title = attributeName, value_MONDO = attributeName_annotated) %>%
#   left_join(create_annotationDF_fromDir(bto_1) %>%
#               rename(study_title = attributeName, value_BTO = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(uberon_1) %>%
#               rename(study_title = attributeName, value_UBERON = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(meddra_1) %>%
#               rename(study_title = attributeName, value_MEDDRA = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(dron_1) %>%
#               rename(study_title = attributeName, value_DRON = attributeName_annotated), by = "study_title")
# 
# df.missing <- create_annotationDF_fromDir(mondo_2) %>% rename(study_title = attributeName, value_MONDO = attributeName_annotated) %>%
#   left_join(create_annotationDF_fromDir(bto_2) %>%
#               rename(study_title = attributeName, value_BTO = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(uberon_2) %>%
#               rename(study_title = attributeName, value_UBERON = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(meddra_2) %>%
#               rename(study_title = attributeName, value_MEDDRA = attributeName_annotated), by = "study_title") %>%
#   left_join(create_annotationDF_fromDir(dron_2) %>%
#               rename(study_title = attributeName, value_DRON = attributeName_annotated), by = "study_title")
# 
# df.study <- df.study %>%
#   bind_rows(df.missing) %>%
#   mutate_all(tolower) %>%
#   distinct()
# 
# write_csv(df.study, "all_study_title_annotations_uncleaned.csv")
```

```{r}
# df.study <- vroom("all_study_title_annotations_uncleaned.csv")
```
# Sample Cleaning

#### Remove duplicate rows
```{r}
df.values <- df.values %>% 
  mutate(entries = apply(., 1, function(x) sum(!is.na(x)))) %>% 
  group_by(value) %>% 
  mutate(n = length(value)) %>% 
  arrange(desc(entries), .by_group = T) %>% 
  slice(1) %>% 
  select(-entries, -n) %>% 
  ungroup()
```


#### Disease 
```{r}
# Remove non-meaningful disease annotations
df.values <- df.values %>% mutate_at(.vars = c("value_MONDO", "value_MEDDRA"), function(x) ifelse(x %in% non_meaningfull_disease_annotations, NA, x))

# Remove obviously incorrect disease annotations
df.values <- df.values %>% mutate_at(.vars = c("value_MONDO", "value_MEDDRA"), function(x) ifelse(x %in% real_study_diseases, x, NA))
df.values <- df.values %>% mutate(disease = pmap_chr(list(value_MONDO, value_MEDDRA), disease_recommender))
```


#### Tissue 
```{r tissueAnnotation}
df.values <- df.values %>% 
  mutate(tissue = pmap_chr(
    list(value_BTO, value_UBERON, value_CLO, value_MONDO, value_MEDDRA), 
    tissue_recommender))
```

#### Cell type 
```{r}
# Remove non-latin characters (CJK Unified Ideographs, CJK Symbols and Punctuation)
# https://en.wikipedia.org/wiki/Unicode_block
df.values <- df.values  %>%
    mutate(value_CLO = gsub("[\U4E00-\U9FFF\U3000-\U303F]", "", value_CLO)) %>%  
    mutate(value_CLO = ifelse(value_CLO =="", NA, value_CLO)) %>%  # Convert "" resulting from unicode correction to NA
    mutate(value_CLO = trimws(value_CLO))

# Apply cell_recommender
df.values <- df.values %>% 
    mutate(cell = pmap_chr(
      list(value_BTO, value_UBERON, value_CLO, value_MONDO, value_MEDDRA), 
      cell_recommender))

# Apply immune lineage recommendations
immune_lineage_key <- make_immune_lineage_key()
df.values <- df.values %>% 
  filter(grepl(paste(immune_lineage_key$pattern, collapse = "|"), value))  %>% 
  mutate(immune_lineage = map_chr(value, function(x) immune_lineage_match(x, immune_lineage_key))) %>% 
  select(value, immune_lineage) %>% 
  left_join(x=df.values,y=., by = "value") %>% 
  mutate(immune_lineage = str_remove(immune_lineage, "_monocyte|monocyte_"))

# Apply immune subset recommendations
immune_subset_key <- make_immune_subset_key()
immune_name_order_key <- make_immune_name_order()
df.values <- df.values %>% 
  filter(!is.na(immune_lineage))  %>%
  mutate(immune_subset = map2_chr(value, immune_lineage, 
                        function(x, y) immune_subset_match(x, y, immune_subset_key, immune_name_order_key))) %>%
  mutate(immune_subset = ifelse(immune_subset == "", immune_lineage, immune_subset)) %>% 
  select(value, immune_lineage, immune_subset) %>%
  left_join(x=df.values ,y=., by = c("value", "immune_lineage"))

# Consolidate cell and immune cell recommendation
df.values <- df.values %>% 
  mutate(cell = ifelse(
    !is.na(immune_subset),
    immune_subset,
    cell
  )) %>% 
  select(-immune_lineage, -immune_subset)

# Apply stem cell recommendations
stem_subset_key <- make_stem_subset_key()
df.values <- df.values %>% mutate(stem_cell = ifelse(grepl(stem_pattern, value), "stem cell", NA)) 
df.values <- df.values %>% 
  select(value, cell, stem_cell) %>% 
  filter(stem_cell == "stem cell") %>% 
  # filter(grepl(paste(stem_subset_key$pattern, collapse = "|"), value))  %>% 
  mutate(stem_subset = map_chr(value, function(x) stem_subset_match(x, stem_subset_key)))%>% 
  select(value, stem_subset) %>% 
  left_join(x=df.values,y=., by = "value") %>% 
  # Remove stem cell derived cells
  mutate(stem_cell = ifelse(grepl("derived|differentiate|from", value), NA, stem_cell)) %>% 
  mutate(stem_subset = ifelse(grepl("derived|differentiate|from", value), NA, stem_subset)) %>% 
  # Add back derived mesenchymal stem cells
  mutate(stem_cell = ifelse(grepl("mesenchym", value), "stem cell", stem_cell)) %>%
  mutate(stem_subset = ifelse(grepl("mesenchym", value), "mesenchymal stem cell", stem_subset)) %>%
  # Consolidate cell. stem_cell, and stem_subset
  mutate(cell = ifelse(!is.na(stem_subset), stem_subset, ifelse(!is.na(stem_cell), stem_cell, cell))) %>% 
  select(-stem_cell, -stem_subset)
```

#### Cell lines

```{r}
df.values <- df.values %>% 
  mutate(is_cell_line = ifelse(!is.na(cell),
              (cell %in% cell.lines) | 
                grepl("[ _\\-]line[ s,._\\:\\)]|[ \\-]line$", value),
              NA)
  )
```

#### Healthy samples
```{r}
df.values <- df.values %>% mutate(is_healthy = ifelse(
    grepl("normal|health", value) & !grepl(" and ", value),
      # values with "and" tend to represent pooled disease and control samples
    TRUE,
    NA
    )
  )
```

#### Control Samples
# Add wt?
```{r}
df.values <- df.values %>% mutate(is_control = ifelse(
    grepl("control|ctrl|placebo", value),
    TRUE,
    NA
    )
  )
```

#### Treatment
```{r}
df.values <- df.values %>% 
  # Correct value_DRON with meaningful treatments
  mutate(match_DRON = ifelse(value_DRON %in% meaningful_treatments, value_DRON, NA)) %>% 
  # Regex using meaningful_treatments to capture non-value_DRON successful titles
  # Will also allow more than one match
  mutate(match_REGEX = unlist(map(value, function(x) {
    x <- unlist(strsplit(x, " |_|-|/"))
    x <- paste(unique(x[x %in% meaningful_treatments]), collapse = "_____")
    ifelse(x == "" | is.na(x), NA, x)
  }))) %>% 
  # Combine, giving preference to match_REGEX > match_DRON
  mutate(treatment = ifelse(is.na(match_REGEX), match_DRON, match_REGEX)) %>%
  select(-match_DRON, -match_REGEX, treatment)
```

#### Ethnicity
```{r}
ethnicity_key <- tibble(values = 
  (map(list.ethnicities, function(x) paste(sort(x), collapse="_____")) %>% 
      map2(names(.), .,
         function(x, y){
           tibble(
             value = x, 
             ethnicity = y)
         }
    ))
  ) %>% 
  pull(values) %>%
  reduce(bind_rows) %>% 
  filter(ethnicity != "unspecified")

df.values <- df.values %>% 
  left_join(ethnicity_key, by = "value")
```

#### Sex 
```{r}
sex_key <- make_sex_key()

df.values <- df.values %>% 
  filter(grepl(paste(sex_key$pattern, collapse = "|"), value))  %>% 
  mutate(sex = map_chr(value, function(x) sex_match(x, sex_key))) %>% 
  select(value, sex) %>% 
  left_join(x=df.values,y=., by = "value") %>% 
  mutate(sex = ifelse(sex == "female_male", "pooled", sex))
```
#### Single cell
```{r}
# x <- df.values %>% 
#   mutate(sc_study = ifelse(grepl(sc_value_pattern, value) &
#     
#   ))
```



#### Drop columns

```{r}
df.values <- df.values %>% 
  select(!contains(c("MONDO", "BTO", "UBERON", "MEDDRA", "CLO", "DRON"))) %>% 
  rename_at(vars(-value), paste0, ".REC")
```

# Age annotation

Some notes on age
- It may be safe to assume that all "unspecified" units are years
- Almost all week units appear to refer to gestation age, which could be either for a pregnant mother or the fetus
- Days seem more likely to refer to days of treatment etc. rather than days of life

```{r}
# Age cleaning
age.tags <- c("age", "donor_age", "age group", "dev_stage", "patient age", "average age", 
              "age of infant", "source subject age", "developmental stage", "age at draw", 
              "age at sampling", "subject age", "sampling age", "age at biopsy", 
              "age at collection", "age of patient", "age group",  
              "median age", "age of donor", "age of human", "age of the fetus",
              "age_category", "age of subjects", "agegroup", "donors age group", "age range")

df.age <- read_csv("./all_tag_values.csv") %>% filter(tag %in% age.tags) %>% 
  mutate(age_unit = map_chr(value, unit.finder)) %>% 
  mutate(test = map2_chr(age_unit, value, time.cleaner)) %>% 
  select(-value, -age_unit) %>% 
  separate(test, sep = "_____", into = c("value", "age_unit", "age_value")) %>% 
  mutate_all(function(x) ifelse(x == "NA", NA, x))

# Year cleaning
year.tags <- c("age_years", "age in years","age_yrs", "patient_age_years")

df.years <- read_csv("./all_tag_values.csv") %>% filter(tag %in% year.tags) %>% 
  mutate(age_unit = "year") %>% 
  mutate(test = map2_chr(age_unit, value, time.cleaner)) %>% 
  select(-value, -age_unit) %>% 
  separate(test, sep = "_____", into = c("value", "age_unit", "age_value")) %>% 
  mutate_all(function(x) ifelse(x == "NA", NA, x))


# Month cleaning
month.tags <- c("age(months)")

df.months <-read_csv("./all_tag_values.csv") %>% filter(tag %in% month.tags) %>% 
  mutate(age_unit = "month") %>% 
  mutate(test = map2_chr(age_unit, value, time.cleaner)) %>% 
  select(-value, -age_unit) %>% 
  separate(test, sep = "_____", into = c("value", "age_unit", "age_value")) %>% 
  mutate_all(function(x) ifelse(x == "NA", NA, x))


# Day cleaning
# NOTE: Compared to years/months, day more likely to refer to non-age related variables (e.g. days treatment, etc.)
day.tags <- c("day of life")

df.days <- read_csv("./all_tag_values.csv") %>% filter(tag %in% day.tags) %>% 
  mutate(age_unit = "day") %>% 
  mutate(test = map2_chr(age_unit, value, time.cleaner)) %>% 
  select(-value, -age_unit) %>% 
  separate(test, sep = "_____", into = c("value", "age_unit", "age_value")) %>% 
  mutate_all(function(x) ifelse(x == "NA", NA, x))

# Gestational week cleaning
gw.tags <- c("gestational week", "post conception week", "weeks of gestation", "gestation_age_sample_collection_weeks", "gestational weeks")

df.gw <-  read_csv("./all_tag_values.csv") %>% filter(tag %in% gw.tags) %>% 
  mutate(age_unit = "week - gestation") %>% 
  mutate(test = map2_chr(age_unit, value, time.cleaner)) %>% 
  select(-value, -age_unit) %>% 
  separate(test, sep = "_____", into = c("value", "age_unit", "age_value")) %>% 
  mutate_all(function(x) ifelse(x == "NA", NA, x))

# Combine
df.age <- bind_rows(df.age, df.months, df.days, df.gw)
rm(df.months, df.days, df.gw, df.years)

# Gestational correction
df.age <- df.age %>% 
  mutate(age_unit = 
  ifelse((age_unit == "category" | grepl("gestation", age_unit)), age_unit, #exclude category unit and units that already note gestation
    ifelse(
    grepl("gest|gw|embry|concep|menst|fet", value),
    paste0(age_unit, " - gestation"),
    age_unit
    )
  ))

df.age <- df.age %>% 
  rename_at(vars(-value, -tag), paste0, ".REC")
```

#### Create tag_value compilation
```{r}
# df.tag_value <- read_csv("all_tag_values.csv") %>% 
#   left_join(df.values, by = "value") %>% 
#   left_join(df.age, by = c("tag", "value")) %>% 
#   mutate(tag = paste0(tag, ".ORIG")) #necessary to prevent tag name conflicts with already existing column names
```

#### Identify single cell samples

```{r}
# df.tag_value <- df.tag_value %>% 
#   # Mark all samples with a single cell value
#   mutate(sc_sample = ifelse(grepl(sc_value_pattern, value), TRUE, FALSE)) %>% 
#   # Mark all samples with a single cell tag
#   mutate(sc_sample = ifelse(grepl(sc_value_pattern, tag), TRUE, sc_sample)) %>% 
#   # Falsify samples that have single cell tag but false value
#   mutate(sc_sample = ifelse(sc_sample == TRUE & grepl(not_sc_pattern, value), FALSE, sc_sample))

```


# Study annotation

#### Disease
```{r}
# Remove non-meaningful disease annotations
df.studies <- df.studies %>% mutate_at(.vars = c("value_MONDO", "value_MEDDRA"), function(x) ifelse(x %in% non_meaningfull_disease_annotations, NA, x))

# Remove obviously incorrect disease annotations
df.studies <- df.studies %>% mutate_at(.vars = c("value_MONDO", "value_MEDDRA"), function(x) ifelse(x %in% real_study_diseases, x, NA))

# Applying disease_recommender
df.studies <- df.studies %>% mutate(disease = pmap_chr(list(value_MONDO, value_MEDDRA), disease_recommender))
```

```{r}
## Capturing un-annotated titles with an additional regex search

# Creating a regex key out of the real_study_diseases list
study_key <- tibble(id = real_study_diseases, pattern = real_study_diseases) %>% 
  filter(nchar(pattern) >4) %>% 
          # exclude patterns < 5 characters which are usually abbreviations 
          # resulting in off target hits
  filter(!(pattern %in% c(
    "stemi" # exclude 'stemi' which will be found in any "systemic" occurence
  )))

# When multiple matches, will only keep the longest one
# Prevents results like 'b cell lymphoblastic leukemia_lymphoblastic leukemia_leukemia"
find_max <- function(x) {x[nchar(x) ==max(nchar(x))]}

# Pattern match updated to include find_max function
study_pattern_match <- function(input, key){
  key %>%
    mutate(match = map2_chr(pattern, id, 
                            function(pattern, id) ifelse(grepl(pattern, input), id, NA))) %>%
    filter(!is.na(match)) %>%
    pull(match) %>%
    find_max() %>% 
    paste(., collapse = "_") %>%
    ifelse(.=="", NA, .)
}

df.studies <- df.studies %>% 
  filter(is.na(disease)) %>% 
  filter(grepl(paste(study_key$pattern, collapse = "|"), study_title))  %>% 
  mutate(regex = map_chr(study_title, function(x) study_pattern_match(x, study_key))) %>% 
  select(study_title, regex) %>% 
  left_join(x=df.studies,y=., by = "study_title") %>% 
  mutate(study_disease.REC = 
           ifelse(is.na(disease), regex, disease)) %>% 
  select(-disease, -regex)
  
```

### Treatments
```{r}
df.studies <- df.studies %>% 
  # Correct value_DRON with meaningful treatments
  mutate(match_DRON = ifelse(value_DRON %in% meaningful_treatments, value_DRON, NA)) %>% 
  # Regex using meaningful_treatments to capture non-value_DRON successful titles
  # Will also allow more than one match
  mutate(match_REGEX = unlist(map(study_title, function(x) {
    x <- unlist(strsplit(x, " |_|-|/"))
    x <- paste(unique(x[x %in% meaningful_treatments]), collapse = "_____")
    ifelse(x == "" | is.na(x), NA, x)
  }))) %>% 
  # Combine, giving preference to match_REGEX > match_DRON
  mutate(study_treatment.REC = ifelse(is.na(match_REGEX), match_DRON, match_REGEX)) %>% 
  select(-match_DRON, -match_REGEX) %>% 
  select(!contains(c("MONDO", "BTO", "UBERON", "MEDDRA", "CLO", "DRON")))
```

### Single cell
```{r}
# df.studies <- df.studies %>% 
#   mutate(sc_study = ifelse(grepl(sc_value_pattern, study_title), TRUE, FALSE)) %>% 
#   select(!contains(c("MONDO", "BTO", "UBERON", "MEDDRA", "CLO", "DRON")))
  
```


# Final annotation compilation

### Compile original metadata, value annotation, and study annotation
```{r}
df.final <- df.original_metadata  %>%
  select(sample_accession, study_title, tag, value) %>%
  mutate_all(tolower) %>%
  distinct() %>%
  left_join(df.studies, by = "study_title") %>%
  left_join(df.values, by = "value") %>%
  left_join(df.age, by = c("tag", "value")) 

# %>% 
  # mutate(tag = paste0(tag, ".ORIG")) #necessary to prevent tag name conflicts with already existing column names

# df.final <- df.original_metadata  %>%  
#   select(sample_accession, study_title, tag, value) %>% 
#   mutate_all(tolower) %>%
#   mutate(tag = paste0(tag, ".ORIG")) %>%  #necessary to prevent tag name conflicts with already existing column names
#   distinct() %>% 
#   left_join(df.studies, by = "study_title") %>% 
#   left_join(df.tag_value, by = c("tag", "value"))
```

### Single cell study identification
```{r}
# x <- df.final %>% 
#   select(sample_accession, study_title, tag, value) %>% 
#   # Mark all samples with a single cell value
#   mutate(sc_sample = ifelse(grepl(sc_pattern, value), TRUE, FALSE)) %>%
#   # Mark all samples with a single cell tag
#   mutate(sc_sample = ifelse(grepl(sc_pattern, tag), TRUE, sc_sample)) %>%
#   # Falsify samples that have single cell tag but false value
#   mutate(sc_sample = ifelse(sc_sample == TRUE & grepl(not_sc_pattern, value), FALSE, sc_sample)) %>% 
#   # Mark all samples from a study with single cell data
#   mutate(sc_study = ifelse(grepl(sc_pattern, study_title), TRUE, FALSE))

df.single_cell <- df.final %>% 
  select(study_title, tag, value) %>% 
  distinct() %>% 
  # Mark all samples with a single cell value
  mutate(sc_sample = ifelse(grepl(sc_pattern, value), TRUE, FALSE)) %>%
  # Mark all samples with a single cell tag
  mutate(sc_sample = ifelse(grepl(sc_pattern, tag), TRUE, sc_sample)) %>%
  # Falsify samples that have single cell tag but false value
  mutate(sc_sample = ifelse(sc_sample == TRUE & grepl(not_sc_pattern, value), FALSE, sc_sample)) %>% 
  # Mark all samples from a study with single cell data
  mutate(sc_study = ifelse(grepl(sc_pattern, study_title), TRUE, FALSE)) %>% 
  # Compile sc_study and sc_sample
  mutate(is_sc = sc_sample | sc_study) %>% select(-sc_sample, -sc_study) %>%
  # Mark study as single cell is any sample in study is marked as single cell
  group_by(study_title) %>% 
  summarise(is_sc_study.REC = any(is_sc))

# x %>% count(is_sc_study)

```

### Convert to long format
```{r}
# Fold recommended columns into the overall long format
# Involves pivot_longer recommended columns and row binding them to original values
df.final <- df.final %>%
  mutate(tag = paste0(tag, ".ORIG")) %>%  #necessary to prevent tag name conflicts with already existing column names
  mutate_all(as.character) 

df.final <- df.final %>% 
  select(-tag, -value) %>% 
  pivot_longer(
    col = contains("REC"),
    names_to = "tag", 
    values_to = "value") %>% 
  bind_rows(df.final) %>% 
  select(sample_accession, sample_accession, tag, value) %>%
  filter(!is.na(value)) 
```

### Combining all metadata annotation
```{r}
# Join full key to original metadata and split into list column based on study
# df.final <- df.original_metadata %>%  
df.final <- df.original_metadata %>%  
  select(-tag, -value) %>%
  distinct() %>%
  left_join(df.final, by = "sample_accession") %>%
  left_join(df.single_cell, by = "study_title") %>% 
  mutate(grouping = study_accession) %>%
  group_by(grouping) %>%
  nest() %>%
  ungroup()

# saveRDS(y, "temp_compiled_df.RDS")
```

```{r, eval=F}
# # Write studies to long form  
# df.final %>% head(10) %>% mutate(data = map2(data, grouping, function(data, grouping) write_csv(data, paste0(grouping,"_annotatedMetadata_long.csv"))))

# Write studies to wide form
df.wide <- df.final %>%
  mutate(data = map(data, function(x)
    {
    pivot_wider(distinct(x),
                names_from = tag,
                values_from = value,
                values_fn = list(value = function(y) paste0(y, collapse = "_____"))
                # ,
                # names_repair = "unique"
                ) %>% 
      # Order each study DF with base->recommendedVariable->originalVariables
      relocate(contains("ORIG"), contains("REC"), .after = last_col()) %>% 
      select(-library_source:-instrument_model, -study_type, -organism, -taxon_id)
    }
  )
  )


# saveRDS(df.wide, "compiled_sra_annotation.RDS")
saveRDS(df.wide, 
        paste0("./compiled_full_sra_annotation_",
               format(Sys.time(), "%Y_%m_%d_%H%M%S"),
               ".RDS"))

# set.seed(10)
# df.wide %>% sample_n(20) %>% mutate(data = map2(data, grouping, function(data, grouping) write_csv(data, paste0("sra_exported/",grouping,"_annotatedMetadata.csv"))))
```




# Writing individual study data
```{r}
df.wide <- readRDS("../compiled_full_sra_annotation_2020_07_19_122311.RDS")
```


#### Comparing new annotations to same files previously sent to group
```{r}
tibble(files = list.files("../sra_exported/")) %>% 
  filter(grepl("annotated", files)) %>% 
  mutate(files = str_remove(files, "_annotatedMetadata.csv")) %>% 
  pull(files) %>% dput()
  
```
```{r}
hong <- c("srp053046", "srp058773", "srp063875", "srp064652", "srp071876")
mich <- c("srp073347", "srp076842", "srp078500", "srp079850", "srp089727")
purv <- c("erp020462", "erp024526", "srp011369", "srp042112", "srp045481")
all_studies <- c(purv, mich, hong)
```

```{r}
df.wide <- readRDS("../compiled_full_sra_annotation_2020_07_19_122311.RDS")
to_export <- df.wide %>% filter(grouping %in% all_studies)
to_export %>% 
  mutate(data = map2(data, grouping, 
                     function(data, grouping) 
                       write_csv(data, paste0("../sra_exported/v2/",grouping,"_annotatedMetadata.csv"))))

system("cd ../sra_exported/v2/;rm ../michele_sra_new.zip;ls|tail -n 5|zip ../michele_sra_new.zip -@")

```
```{r}
old <- readRDS("../compiled_sra_annotation.RDS")
old %>% mutate(n = reduce(map(data, nrow), rbind)) %>% summarise(sum(n))
```




